\textbf{ Spectrophotometric parallaxes with linear models:\\ Accurate distances for luminous red-giant stars }

### 2019-06-02 paper/response_to_referee.txt

    The paper presents a method to estimate spectrophotometric parallaxes
    based only on input data, without adoption of physically motivated
    models and priors. The paper is written in an way that is strongly
    focused on the method itself, but leaves several complementary aspects
    that are certainly important to assess the reliability of the inferred
    parallaxes and, most importantly, distances. I recommend its
    publication only after such major issues are addressed.

    Major issues

    1 - The data-driven method described in the paper details out how
    spectrophotometric parallaxes are inferred. But the paper does not
    give the reader a clue about how spectrophotometric distances are
    estimated. The authors make it clear in Section 2 that they never
    invert *observed* parallaxes to estimate distances in their
    method. But are the spectrophotometric distances they infer (shown in
    Figures 5 and 6) just the inverse of the *inferred* parallaxes? If so,
    this should be made explicit in the text, along with some discussion
    about the problems (or lack of) that may arise from this
    inversion. Even though the spectrophotometric parallaxes themselves
    are always positive, a reader would like to be convinced that the
    distances shown in Figures 5 and 6 do not suffer from biasses caused
    by the asymmetric nature of the distances distribution that results
    from a symmetric error distribution in parallaxes (as stated by the
    likelihood assumption of the method). How do these biasses simply go
    away when the method's w(sp) +/- sigma(sp) are used? In particular, a
    negative parallax for many stars may be consistent with their w(sp)
    +/- sigma(sp).

These issues are subtle! We are assuming that the GAIA-MEASURED
PARALLAXES have noise that is symmetric in parallax space. In our
data-driven model, we use a linear combination of features to predict
or estimate a LOG PARALLAX, so if anything we are claiming that our
estimates (not the data, but our estimates based on the data) are
symmetric in log parallax, not parallax. But we aren't even really
claiming that. The issues are subtle. However, this does deserve some
discussion and we have added that to the captions of Figure XXX and
YYY (formerly Figure 5 and 6), and in the last paragraph of the
discussion, for extra-special emphasis.  The short answer is that we
just exponentiated the negative of our log-parallax estimates to make
those figures.

### 2019-06-02 paper/sp.tex
one more item off

Finally, one comment on the use of these spectrophotometric parallaxes.
They are noisy.
[HOGG PLACEHOLDER A]
Precise use of these involves building a likelihood function and performing inference,
just as the precise use of the \gaia\ data presented here required the same.
We recommend making analogous inferences when using these outputs.

>>>>

Finally, one comment on the use of these spectrophotometric parallaxes.
They are noisy.
And they linearly combine a set of features into a logarithmic parallax or
logarithmic distance.
Because the input features have close-to-Gaussian noise (mainly
because they are observed at high signal-to-noise), the linear
combination will have close-to-Gaussian noise.
That is, the noise in the estimated distances or parallaxes ought to be close
to Gaussian in the logarithm.
This is a consideration in the use of these outputs:
Precise use of these involves building a likelihood function and performing inference,
just as the precise use of the \gaia\ data presented here required the same.
Justified likelihood functions are critical to making good inferences.
